Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm} Stagnating State Funding for Higher Education and its Effect on Faculty

### 2024-05-23 text/paper.tex
Added a brief first-stage spending analysis.
Adjusted formatting of some figures, and changed some bad punctuation.

% Page settings
\documentclass[notitlepage,12pt]{article}
\usepackage{fancyvrb, verbatim, listings}                                  % Various for inserting/hosting code
\usepackage{setspace}                                                      % Space between lines (customisable)
\usepackage[margin=2.54cm]{geometry}                                       % Set margins (standard is 2.54cm)
\usepackage[UKenglish,cleanlook]{isodate}                                  % Set default date and date display
\usepackage{fancyhdr} \pagestyle{fancy}                                    % Line the top of the page
\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\microtypecontext{spacing=nonfrench}                                       % Change font and minor spacings to look nicer
\setlength{\parskip}{0.5\baselineskip}                                     % Set default paragraph behaviour: skip a line
%\setlength{\parindent}{0pt}                                                % Set default paragraph behaviour: no indent
%% Packages to use:
% Tables + figures
\usepackage{graphicx}                                                      % control over the import of graphics
\usepackage[table,dvipsnames]{xcolor}                                      % Define tables (with colour control)
\usepackage{float}                                                         % Control over graphics + tables float
\usepackage[justify]{ragged2e}                                             % control over text alignment
\usepackage{booktabs}                                                      % caption graphics + tables (with name in bold)
\usepackage[labelfont=bf]{caption} 
\usepackage{subcaption}
% Maths + numbers
\usepackage{mathtools}                                                     % Various maths functions
\usepackage{amssymb}                                                       % Various maths functions
\usepackage{amsmath}                                                       % Various maths functions
\usepackage{dsfont}                                                        % Various maths functions
\usepackage{centernot}                                                     % center \not usage
\usepackage{siunitx} \sisetup{round-mode=places, round-precision=3}        % Formalise use of units and numbers among text
\DeclareMathOperator{\eps}{\varepsilon}                                    % epsilon short hand shortcut
\renewcommand{\vec}[1]{\boldsymbol{\mathit{#1}}}                           % vector notation shortcut
\newcommand{\mat}[1]{\boldsymbol{\mathit{#1}}}                             % matrix notation shortcut
\renewcommand{\hat}[1]{\widehat{#1}}                                       % Default estimator notation is widehat
\renewcommand{\bar}[1]{\overline{#1}}                                      % Make over bar look nicer
\renewcommand{\tilde}[1]{\widetilde{#1}}                                   % Make over tilde look better
% Citations
\usepackage[longnamesfirst]{natbib}                                        % Citation package, see https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management#Natbib
\usepackage[backref=page]{hyperref}                                        % Allow for links across the text, with colour options
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=magenta, urlcolor=blue}
\def\sectionautorefname~#1\null{Section~#1\null}                           % Fix autoref for sections
\def\subsectionautorefname~#1\null{Subsection~#1\null}                     % Fix autoref for sections
\def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for equations
% IMPORTANT: follow style guide here https://github.com/Wookai/paper-tips-and-tricks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title page
% Author
\author{Senan Hogan-Hennessy\thanks{
        For helpful comments I thank
        Levon Barseghyan,
        Francine Blau,
        Romain Campbell,
        Maddalena Davoli,
        Ryan Dycus,
        Ronald Ehrenberg,
        Michael Lovenheim,
        Jake Meyer,
        Douglas Miller,
        and Evan Riehl.
        I thank seminar participants at Cornell University (2023, 2024), the Eastern Economic Association Annual Meetings (2024) for helpful discussion.
        This project's repository, with all materials and anonymised data needed for replication, is available at 
        \url{https://github.com/shoganhennessy/state-funding-faculty}.
        Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
    } \\
    \vspace{0.1cm}
    Economics Department, Cornell University\footnote{
        Address: Uris Hall \#447, Economics Department, Cornell University NY 14853 USA.
    }
}
% Title
\title{Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm}
    Stagnating State Funding for Higher Education and its Effect on Faculty}
\date{This version: \today \\ \vspace{0.1cm}
    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2024.pdf}{Newest version available here.}
}
\rhead{\today}
\lhead{Less Funding, More Lecturers, and Fewer Professors.}
% Begin
\begin{document}
\maketitle
\thispagestyle{empty}
% Abstract
\begin{abstract}
    \noindent
    \input{sections/00-abstract.tex}
\end{abstract}

>>>>

% Page settings
\documentclass[notitlepage,12pt]{article}
\usepackage{fancyvrb, verbatim, listings}                                  % Various for inserting/hosting code
\usepackage{setspace}                                                      % Space between lines (customisable)
\usepackage[margin=2.54cm]{geometry}                                       % Set margins (standard is 2.54cm)
\usepackage[UKenglish,cleanlook]{isodate}                                  % Set default date and date display
\usepackage{fancyhdr} \pagestyle{fancy}                                    % Line the top of the page
\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\microtypecontext{spacing=nonfrench}                                       % Change font and minor spacings to look nicer
\setlength{\parskip}{0.5\baselineskip}                                     % Set default paragraph behaviour: skip a line
%\setlength{\parindent}{0pt}                                                % Set default paragraph behaviour: no indent
%% Packages to use:
% Tables + figures
\usepackage{multirow}
\usepackage{graphicx}                                                      % control over the import of graphics
\usepackage[table,dvipsnames]{xcolor}                                      % Define tables (with colour control)
\usepackage{float}                                                         % Control over graphics + tables float
\usepackage[justify]{ragged2e}                                             % control over text alignment
\usepackage{booktabs}                                                      % caption graphics + tables (with name in bold)
\usepackage[labelfont=bf]{caption} 
\usepackage{subcaption}
% Maths + numbers
\usepackage{mathtools}                                                     % Various maths functions
\usepackage{amssymb}                                                       % Various maths functions
\usepackage{amsmath}                                                       % Various maths functions
\usepackage{dsfont}                                                        % Various maths functions
\usepackage{centernot}                                                     % center \not usage
\usepackage{siunitx} \sisetup{round-mode=places, round-precision=3}        % Formalise use of units and numbers among text
\DeclareMathOperator{\eps}{\varepsilon}                                    % epsilon short hand shortcut
\renewcommand{\vec}[1]{\boldsymbol{\mathit{#1}}}                           % vector notation shortcut
\newcommand{\mat}[1]{\boldsymbol{\mathit{#1}}}                             % matrix notation shortcut
\renewcommand{\hat}[1]{\widehat{#1}}                                       % Default estimator notation is widehat
\renewcommand{\bar}[1]{\overline{#1}}                                      % Make over bar look nicer
\renewcommand{\tilde}[1]{\widetilde{#1}}                                   % Make over tilde look better
% Citations
\usepackage[longnamesfirst]{natbib}                                        % Citation package, see https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management#Natbib
\usepackage[backref=page]{hyperref}                                        % Allow for links across the text, with colour options
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=magenta, urlcolor=blue}
\def\sectionautorefname~#1\null{Section~#1\null}                           % Fix autoref for sections
\def\subsectionautorefname~#1\null{Subsection~#1\null}                     % Fix autoref for sections
\def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for equations
% IMPORTANT: follow style guide here https://github.com/Wookai/paper-tips-and-tricks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title page
% Author
\author{Senan Hogan-Hennessy\thanks{
        For helpful comments I thank
        Levon Barseghyan,
        Francine Blau,
        Romain Campbell,
        Maddalena Davoli,
        Ryan Dycus,
        Ronald Ehrenberg,
        Michael Lovenheim,
        Jake Meyer,
        Douglas Miller,
        and Evan Riehl.
        I thank seminar participants at Cornell University (2023, 2024), and the Eastern Economic Association Annual Meetings (2024) for helpful discussion.
        This project's repository, including all materials and anonymised data needed for replication, is publicly available at 
        \url{https://github.com/shoganhennessy/state-funding-faculty}.
        Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
    } \\
    \vspace{0.1cm}
    Economics Department, Cornell University\footnote{
        Address: Uris Hall \#447, Economics Department, Cornell University NY 14853 USA.
    }
}
% Title
\title{Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm}
    Stagnating State Funding for Higher Education and its Effect on Faculty}
\date{This version: \today \\ \vspace{0.1cm}
    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2024.pdf}{Newest version available here.}
}
\rhead{\today}
\lhead{Less Funding, More Lecturers, and Fewer Professors.}
% Begin
\begin{document}
\maketitle
\thispagestyle{empty}
% Abstract
\begin{abstract}
    \noindent
    \input{sections/00-abstract.tex}
\end{abstract}

### 2024-05-23 text/sections/01-introduction.tex
Adjusted formatting of some figures, and changed some bad punctuation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction section
%\section{Introduction}
%\label{sec:intro}
States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
At the same time, public universities have systematically substituted away from professors towards lecturers, who teach often work on contingent and lower-salaried contracts.
Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.

@@

Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition and wages.
The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
For faculty outcomes, descriptive work measures how US and Canadian faculty salaries are correlated with other factors \citep{bratsberg2003negative,martinello2009faculty,pfeffer1992salary}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
These papers measure changes in state funding and university revenues, but do not measure effects of state funding cuts on faculty composition or salaries, or other outcomes related to university instruction.
My results provide evidence that the long term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction section
%\section{Introduction}
%\label{sec:intro}
States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
At the same time, public universities have systematically substituted away from professors towards lecturers, who often work on contingent and lower-salaried contracts.
Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.

@@

Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition and wages.
The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
For faculty outcomes, descriptive work measures how US and Canadian faculty salaries are correlated with other factors \citep{bratsberg2003negative,martinello2009faculty,pfeffer1992salary,hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
These papers measure changes in state funding and university revenues, but do not measure effects of state funding cuts on faculty composition or salaries, or other outcomes related to university instruction.
My results provide evidence that the long term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.

### 2024-05-23 text/sections/02-data.tex
updated
Added new robustness check tables.
Adjusted formatting of some figures, and changed some bad punctuation.

IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
    IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
}
Data are consistent for 1990--2021,\footnote{
    The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
    Year refers to the calendar year of the spring term --- i.e., 1990 refers to the academic year that ran August 1989 to July 1990.
}
and provide information on university funding, enrolment, and numerous other characteristics.\footnote{
    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990--2017, and manually combine raw National Center for Education Statistics (NCES) data on 2018--2021 for all relevant variables.
    Figures for enrolment and faculty counts come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
}
I restrict analysis to public, four-year, degree-granting institutions, as these institutions have the largest majority of faculty employed and students enrolled.
For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022}, so that these institutions are not considered.
IPEDS reports the count of professors employed by position.\footnote{
    IPEDS gives a mean salary measure for faculty by rank, but these figures have many missing values and disagree with calculated values from other reputable sources.
    Over 40\% of university--year observations are missing in the IPEDS panel data-set, representing an average of 55\% undergraduate enrolment in each year.
    Yearly averages of faculty salaries by the non-missing values do not agree with trends in average professor salary over the sample time period compared to summary statistics provided in \cite{aau2021survey}.
    IPEDS values consistently report that public university professors are paid on average 20\% more than private sector faculty, which disagrees with other sources on the salary gap direction, and IPEDS figures are consistently higher than averages calculated from individual-level IBHED data.
    For these reasons I do not further analyse IPEDS salary data, and use salary data from IBHED to analyse faculty salaries.
    Notably, IPEDS data on faculty counts do not suffer from these problems.
}
This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count\footnote{
    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
} for each faculty position (lecturer, assistant professor, tenured professor, and total faculty), and other university characteristics.
\autoref{tab:ipeds-summary} presents summary statistics for these variables in IPEDS data.

IPEDS provides information at the university level, but only provides aggregated information within a university.
To investigate outcomes for individual faculty members, I integrate individual-level data for all faculty members at all Illinois public universities 2010--2021.
These data allow me to investigate how individual faculty members are affected by funding cuts to their university, investigating outcomes are more granular details than total faculty counts.
IBHED freely provides this information, as the state of Illinois is required to publicly report base salary and benefits for all administrators, faculty members, and instructors employed by each public college or university.\footnote{
    Public Act 96-0266, effective 1 January 2010, is the relevant Illinois law that requires publicly publishing salary data for all public university faculty salary and benefits \citep{illinois-public-act}.
    A pdf copy of Public Act 096-0266 is included in the supplmentary data of this paper.
    This law applies to all nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University (all five campuses), University of Illinois (all four campuses), Western Illinois University.
}
These data provide the basis to build a panel of Illinois public university faculty across years 2010--2021; I define a faculty member as an individual by their first plus last name and university pairing,\footnote{
    Some faculty are listed multiple times in these panel data.
    For example, Professor Alberto Agustin Lopez-Scala of the University of Illinois Chicago has two appointments, one as an adjunct faculty and one as a department director.
    For this observation I take the highest paid position (department director) as the primary appointment and drop the secondary appointment (adjunct faculty).
    In analysis of faculty exit rate, I collapse the appointments into one and only consider faculty exits as having a faculty appointment versus no appointment.
}
and link this database to IPEDS for data on their university employer.
The Illinois sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
Analysis of professors' in their first year on the job focuses on the subset of professors with observed year of hiring 2011--2021 (i.e., after the panel's first year), representing 1,778 professors in 2011, and 9,099 in 2021.

@@

>>>>

IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
    IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
}
Data are consistent for 1990--2021,\footnote{
    The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
    Year refers to the calendar year of the spring term --- i.e., 1990 refers to the academic year that ran August 1989 to July 1990.
}
and provide information on university funding, enrolment, and numerous other characteristics.\footnote{
    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990--2017, and manually combine raw National Center for Education Statistics (NCES) data on 2018--2021 for all relevant variables.
    Figures for enrolment and faculty counts come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
    I also integrate Barron's selectivity index, from the 2009 rankings, to analyse effects for different levels of selectivity \citep{barrons2009}.
}
I restrict analysis to public, four-year, degree-granting institutions, as these institutions have the largest majority of faculty employed and students enrolled.
For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022}, so that these institutions are not considered.
IPEDS reports the count of professors employed by position.\footnote{
    IPEDS gives a mean salary measure for faculty by rank, but these figures have many missing values and disagree with calculated values from other reputable sources.
    Over 40\% of university--year observations are missing in the IPEDS panel data-set, representing an average of 55\% undergraduate enrolment in each year.
    Yearly averages of faculty salaries by the non-missing values do not agree with trends in average professor salary over the sample time period compared to summary statistics provided in \cite{aau2021survey}.
    IPEDS values consistently report that public university professors are paid on average 20\% more than private sector faculty, which disagrees with other sources on the salary gap direction, and IPEDS figures are consistently higher than averages calculated from individual-level IBHED data.
    For these reasons I do not further analyse IPEDS salary data, and use salary data from IBHED to analyse faculty salaries.
    Notably, IPEDS data on faculty counts do not suffer from these problems.
}
This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count for each faculty position (lecturer, assistant professor, tenured professor, and total faculty), and other university characteristics.
\autoref{tab:ipeds-summary} presents summary statistics for these variables in IPEDS data.

IPEDS provides information at the university level, but only provides aggregated information within a university.
To investigate outcomes for individual faculty members, I integrate individual-level data for all faculty members at all Illinois public universities 2010--2021.
These data allow me to investigate how individual faculty members are affected by funding cuts to their university, investigating individual-level professor outcomes unavailable in IPEDS data.
IBHED freely provides this information, as the state of Illinois is required to publicly report base salary and benefits\footnote{
    Real salary is computed by scaling nominal salary and benefits to 2021 dollars by the CPI-U.
} for all administrators, faculty members, and instructors employed by each public college or university.\footnote{
    Public Act 96-0266, effective 1 January 2010, is the relevant Illinois law that requires publicly publishing salary data for all public university faculty salary and benefits \citep{illinois-public-act}.
    A pdf copy of Public Act 096-0266 is included in the supplmentary data of this paper.
    This law applies to all nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University (all five campuses), University of Illinois (all four campuses), Western Illinois University.
}
These data provide the basis to build a panel of Illinois public university faculty across years 2010--2021; I define a faculty member as an individual by their first plus last name and university pairing,\footnote{
    Some faculty are listed multiple times in these panel data.
    For example, Professor Alberto Agustin Lopez-Scala of the University of Illinois Chicago has two appointments in each year 2013--2021, one as an adjunct faculty and one as a department director.
    For this observation I take the highest paid position (department director) as the primary appointment and drop the secondary appointment (adjunct faculty).
    In analysis of faculty exit rate, I collapse the appointments into one and only consider faculty exits as having a faculty appointment versus no appointment.
}
and link this database to IPEDS for data on their university employer.
The Illinois sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
Analysis of professors' in their first year on the job focuses on the subset of professors with observed year of hiring 2011--2021 (i.e., after the panel's first year), representing 1,778 professors in 2011, and 9,099 in 2021.

@@

% Paragraph describing how IL is similar to rest of country.

### 2024-05-23 text/sections/05-results.tex
Adjusted formatting of some figures, and changed some bad punctuation.

\subsection{Robustness Checks}
\label{sec:results-robustness}
It is possible that the identification strategy is not robust to some possible confounders.
For example, it is possible that selective institutions received less funding for other reasons, or states with lower performing institutions or declining populations strategically limited funding for less selective universities.
I address these main concerns with three different analyses.

@@

Secondly, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
This investigates whether results are consistent across different levels of selectivity, and whether less selective universities have unreasonably high estimates for substitution --- as would be the case if selectivity was an unaddressed confounder.
\autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
The first-stage is string in every group of universities, even the most selective group which is only 12 universities, the selective group which is 28, and the rest are 762 universities.
Faculty substitution rates are not very string among the most selective universities, but are very similar between selective universities and unranked universities, showing that even selective public universities exhibit the substitution towards lecturers.

The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.
Private universities do not rely on state funding in the same that public universities do,\footnote{
    Though private universities do receive some state funding, as some states fund for project at private universities and other smaller scale operating costs.
}
so we would expect state funding to not affect their faculty counts.
If the state funding and faculty counts were related among private universities, this would be evidence that the IV model picks up otherwise concurrent trends in higher education, unrelated to stagnating state funding for public universities.
To impute a base share of reliance on state funding, I assign private universities the average reliance on state funding in the base period of universities in the same state with the same Barrons selectivity ranking, and then compute the shift-share instrument from this value.
\autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities; despite this, the IV results are still not significant for any faculty count, showing that state funding is not related to private universities' faculty counts.
The OLS columns show that per student state funding is highly correlated with lecturer and professor counts at private universities.
The IV estimates show that these correlations are not causal relationships, validating the approach for inference on public universities' faculty counts.

>>>>

\subsection{Robustness Checks}
\label{sec:results-robustness}
It is possible that the identification strategy is not robust to some confounding issues or trends.
For example, selective institutions may have received funding cuts for other reasons, or states with lower performing institutions or declining populations strategically cut funding for only less selective universities.
I address concerns for this with three different analyses.

@@

Secondly, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
This investigates whether results are consistent across different levels of selectivity, and whether less selective universities have unreasonably high estimates for substitution --- as would be the case if selectivity was an unaddressed confounder.
\autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
The first-stage is strong in every group of universities, even among the small sample size of the 12 universities in the most selective group, and the 28 universities in the next selective group.
Faculty substitution rates are not very strong among the most selective universities, but are very similar between selective universities and unranked universities, showing that even selective public universities exhibit the substitution towards lecturers.

The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.
Private universities do not rely on state funding in the same that public universities do,\footnote{
    Though private universities do receive some state funding, as some states fund for project at private universities and other smaller scale operating costs.
}
so we would expect state funding to not affect their faculty counts.
If the state funding and faculty counts were related among private universities, this would be evidence that the IV model picks up otherwise concurrent trends in higher education, unrelated to stagnating state funding for public universities.
To impute a base share of reliance on state funding, I assign private universities the average reliance on state funding in the base period of universities in the same state with the same Barrons selectivity ranking, and then compute the shift-share instrument from this value.
\autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities; despite this, the IV results are still not significant for any faculty count, showing that state funding is not related to private universities' faculty counts.
The OLS columns show that imputed state funding is highly correlated with lecturer and professor counts at private universities.
The IV estimates are not distinguishable from zero, showing that OLS correlations are not causal relationships, and validating the IV approach for inference on universities' faculty counts.
